% Proposing a table of examples to include

\subsection{Correctly handled examples}

\fracasex{252}{\answer yes
\item	Since 1992 ITEL has been in Birmingham.
\item	It is now 1996.
\hyp    Itel was in Birmingham in 1993.

} \fracasex{253	}{\answer unknown
\item	ITEL has developed a new editor since 1992.
\item	It is now 1996.
\hyp    ITEL developed a new editor in 1993.


} \fracasex{254	}{\answer unknown
\item	ITEL has expanded since 1992.
\item	It is now 1996.
\hyp    ITEL expanded in 1993.


} \fracasex{258	}{\answer no
\item	In March 1993 APCOM founded ITEL.
\hyp    ITEL existed in 1992.



} \fracasex{260	}{\answer yes
\item	Yesterday APCOM signed the contract.
\item	Today is Saturday, July 14th.
\hyp    APCOM signed the contract Friday, 13th.


} \fracasex{261	}{\answer yes
\item	Smith left before Jones left.
\item	Jones left before Anderson left.
\hyp    Smith left before Anderson left.


} \fracasex{262	}{\answer yes
\item	Smith left after Jones left.
\item	Jones left after Anderson left.
\hyp    Smith left after Anderson left.

} \fracasex{264	}{\answer yes
\item	Smith left.
\item	Jones left.
\item	Smith left before Jones left.
\hyp    Jones left after Smith left.
} \fracasex{265	}{\answer yes
\item	Smith left.
\item	Jones left.
\item	Smith left after Jones left.
\hyp    Jones left before Smith left.

} \fracasex{266	}{\answer yes
\item	Smith left.
\item	Jones left.
\item	Jones left before Smith left.
\hyp    Smith left after Jones left.
} \fracasex{267	}{\answer yes
\item	Jones revised the contract.
\item	Smith revised the contract.
\item	Jones revised the contract before Smith did.
\hyp    Smith revised the contract after Jones did.
} \fracasex{268	}{\answer yes
\item	Jones revised the contract.
\item	Smith revised the contract.
\item	Jones revised the contract after Smith did.
\hyp    Smith revised the contract before Jones did.


} \fracasex{269	}{\answer unknown
\item	Smith swam.
\item	Jones swam.
\item	Smith swam before Jones swam.
\hyp    Jones swam after Smith swam.

} \fracasex{270	}{\answer yes
\item	Smith swam to the shore.
\item	Jones swam to the shore.
\item	Smith swam to the shore before Jones swam to the shore.
\hyp    Jones swam to the shore after Smith swam to the shore.

} \fracasex{273	}{\answer unknown
\item	Smith was writing a report.
\item	Jones was writing a report.
\item	Smith was writing a report before Jones was writing a report.
\hyp    Jones was writing a report after Smith was writing a report.

} \fracasex{274	}{\answer unknown
\item	Smith was writing a report.
\item	Jones was writing a report.
\item	Smith was writing a report after Jones was writing a report.
\hyp    Jones was writing a report before Smith was writing a report.


} \fracasex{277	}{\answer unknown
\item	Smith lived in Birmingham in 1991.
\hyp    Smith lived in Birmingham in 1992.


} \fracasex{279	}{\answer no
\item	Smith wrote a novel in 1991.
\hyp    Smith wrote it in 1992.

} \fracasex{280	}{\answer unknown
\item	Smith wrote a novel in 1991.
\hyp    Smith wrote a novel in 1992.

} \fracasex{281	}{\answer unknown
\item	Smith was running a business in 1991.
\hyp    Smith was running it in 1992.

} \fracasex{282	}{\answer no
\item	Smith discovered a new species in 1991.
\hyp    Smith discovered it in 1992.

} \fracasex{283	}{\answer unknown
\item	Smith discovered a new species in 1991.
\hyp    Smith discovered a new species in 1992.

} \fracasex{284	}{\answer yes
\item	Smith wrote a report in two hours.
\item	Smith started writing the report at 8 am.
\hyp    Smith had finished writing the report by 11 am.

} \fracasex{287	}{\answer unknown
\item	Smith wrote a report in two hours.
\hyp    Smith wrote a report in one hour.
} \fracasex{288	}{\answer yes
\item	Smith wrote a report in two hours.
\hyp    Smith wrote a report.
} \fracasex{289	}{\answer no
\item	Smith discovered a new species in two hours.
\hyp    Smith spent two hours discovering the new species.
Why 	Achievements are typically (more or less) instantaneous
} \fracasex{290	}{\answer yes
\item	Smith discovered a new species in two hours.
\hyp    Smith discovered a new species.
} \fracasex{291	}{\answer yes **
\item	Smith discovered many new species in two hours.
\hyp    Smith spent two hours discovering new species.
A 	?Yes
Why 	Repeated achievement can last two hours.
} \fracasex{292	}{\answer unknown
\item	Smith was running his own business in two years.
\hyp    Smith spent two years running his own business.
Why 	Premise refers to time taken to inception of activity, not duration of activity.
} \fracasex{293	}{\answer unknown
\item	Smith was running his own business in two years.
\hyp    Smith spent more than two years running his own business.
Why 	Cf. similar inference for accomplishment, (286)
} \fracasex{295	}{\answer unknown
\item	In two years Smith owned a chain of businesses.
\hyp    Smith owned a chain of business for two years.
Why 	States behave like activities.
Note 	Sic: the original did have "chain of business" in the question.
} \fracasex{297	}{\answer yes
\item	In two years Smith owned a chain of businesses.
\hyp    Smith owned a chain of business.
Note 	Sic: the original did have "chain of business" in the question.
} \fracasex{298	}{\answer yes
\item	Smith lived in Birmingham for two years.
\hyp    Smith lived in Birmingham for a year.
Why 	State
} \fracasex{299	}{\answer no
\item	Smith lived in Birmingham for two years.
\hyp    Smith lived in Birmingham for exactly a year.
} \fracasex{300	}{\answer yes
\item	Smith lived in Birmingham for two years.
\hyp    Smith lived in Birmingham.
} \fracasex{301	}{\answer yes
\item	Smith ran his own business for two years.
\hyp    Smith ran his own business for a year.
Why 	Activity
} \fracasex{302	}{\answer yes
\item	Smith ran his own business for two years.
\hyp    Smith ran his own business.
} \fracasex{303	}{\answer yes
\item	Smith wrote a report for two hours.
\hyp    Smith wrote a report for an hour.
Why 	Accomplishment
} \fracasex{304	}{\answer unknown
\item	Smith wrote a report for two hours.
\hyp    Smith wrote a report.
Why 	He may not have finished it

} \fracasex{306	}{\answer yes
\item	Smith discovered new species for two years.
\hyp    Smith discovered new species.


} \fracasex{307	}{\answer yes
\item	In 1994 ITEL sent a progress report every month.
\hyp    ITEL sent a progress report in July 1994.


} \fracasex{311	}{\answer yes
\item	Smith had left the house at a quarter past five.
\item	Then she took a taxi to the station.
\hyp    Smith left the house before she took a taxi to the station.
} \fracasex{312	}{\answer yes
\item	ITEL always delivers reports late.
\item	In 1993 ITEL delivered reports.
\hyp    ITEL delivered reports late in 1993.
Note 	Sic: the original did have "delivered" in the question.
} \fracasex{313	}{\answer no
\item	ITEL never delivers reports late.
\item	In 1993 ITEL delivered reports.
\hyp    ITEL delivered reports late in 1993.
} \fracasex{314	}{\answer yes
\item	Smith arrived in Paris on the 5th of May, 1995.
\item	Today is the 15th of May, 1995.
\item	She is still in Paris.
\hyp    Smith was in Paris on the 7th of May, 1995.


} \fracasex{320	}{\answer yes
\item	When Jones got his job at the CIA, he knew that he would never be allowed to write his memoirs.
\hyp    It is the case that Jones is not and will never be allowed to write his memoirs.
}

\subsection{Incorrectly handled examples}

We list below a number of selected examples from the temporal section
of the FraCaS testsuite where our semantics fails to list the correct
inference. We list the reason why it fails under the item \textbf{F}.

{

} \fracasex{259	}{\answer yes
\item	The conference started on July 4th, 1994.
\item	It lasted 2 days.
\hyp    The conference was over on July 8th, 1994.
\failsf Incorrect lexical semantics for ``last $n$ days''.
} \fracasex{263	}{\answer unknown
\item	Smith was present after Jones left.
\item	Jones left after Anderson was present.
\hyp    Smith was present after Anderson was present.
\failsf Copula do not get timespans

} \fracasex{278	}{\answer no
\item	Smith wrote his first novel in 1991.
\hyp    Smith wrote his first novel in 1992.
\failsf Unicity of ordinals not implemented

} \fracasex{285	}{\answer unknown
\item	Smith wrote a report in two hours.
\hyp    Smith spent two hours writing the report.
\failsf Temporal semanticts of ``spend'' incorrect

} \fracasex{286	}{\answer no
\item	Smith wrote a report in two hours.
\hyp    Smith spent more than two hours writing the report.
\failsf Temporal semanticts of ``spend'' incorrect

} \fracasex{294	}{\answer yes
\item	Smith was running his own business in two years.
\hyp    Smith ran his own business.
\failsf Parsing error for ``in two years''

} \fracasex{296	}{\answer unknown
\item	In two years Smith owned a chain of businesses.
\hyp    Smith owned a chain of business for more than two years.
\failsf Parsing error for ``in two years''

} \fracasex{315	}{\answer yes
\item	When Smith arrived in Katmandu she had been travelling for three days.
\hyp    Smith had been travelling the day before she arrived in Katmandu.
\failsf Incorrect idiomatic semantics for ``the day before''

} \fracasex{316	}{\answer yes
\item	Jones graduated in March and has been employed ever since.
\item	Jones has been unemployed in the past.
\hyp    Jones was unemployed at some time before he graduated.
\failsf Incorrect idiomatic semantics


} \fracasex{317	}{\answer yes
\item	Every representative has read this report.
\item	No two representatives have read it at the same time.
\item	No representative took less than half a day to read the report.
\item	There are sixteen representatives.
\hyp    It took the representatives more than a week to read the report.
\failsf Counting semantics not implemented


} \fracasex{318	}{\answer no
\item	While Jones was updating the program, Mary came in and told him about the board meeting.
\item	She finished before he did.
\hyp    Mary's story lasted as long as Jones's updating the program.
\failsf Incorrect interaction with anaphora


} \fracasex{321	}{\answer yes
\item	Smith has been to Florence twice in the past.
\item	Smith will go to Florence twice in the coming year.
\hyp    Two years from now Smith will have been to Florence at least four times.
\failsf Counting semantics not implemented

} \fracasex{323	}{\answer yes
\item	No one gambling seriously stops until he is broke.
\item	No one can gamble when he is broke.

\hyp    Everyone who starts gambling seriously stops the moment he is broke.
\failsf Idiomatic semantics of ``the moment'' not supported
}
